---
title: "Build a model"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(tibble.print_min = 5)
```

Get started with building a model in this R Markdown document that accompanies [Build a model](https://www.tidymodels.org/start/models/) tidymodels start article.

If you ever get lost, you can visit the links provided next to section headers to see the accompanying section in the online article.

Take advantage of the RStudio IDE and use "Run All Chunks Above" or "Run Current Chunk" buttons to easily execute code chunks. If you have been running other tidymodels articles in this project, restart R before working on this article so you don't run out of memory on RStudio Cloud.


## [Introduction](https://www.tidymodels.org/start/models/#intro)

Load necessary packages:

```{r}
library(tidymodels)  # for the parsnip package, along with the rest of tidymodels

# Helper packages
library(readr)       # for importing data
library(broom.mixed) # for converting bayesian models to tidy tibbles
```


## [The Sea Urchins Data](https://www.tidymodels.org/start/models/#data)


```{r}
urchins <-
  # Data were assembled for a tutorial 
  # at https://www.flutterbys.com.au/stats/tut/tut7.5a.html
  read_csv("https://tidymodels.org/start/models/urchins.csv") %>% 
  # Change the names to be a little more verbose
  setNames(c("food_regime", "initial_volume", "width")) %>% 
  # Factors are very helpful for modeling, so we convert one column
  mutate(food_regime = factor(food_regime, levels = c("Initial", "Low", "High")))
```

Look at the data:

```{r}
urchins
```

Plot the data: 

```{r}
ggplot(urchins,
       aes(x = initial_volume, 
           y = width, 
           group = food_regime, 
           col = food_regime)) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE) +
  scale_color_viridis_d(option = "plasma", end = .7)
```

## [Build and fit a model](https://www.tidymodels.org/start/models/#build-model)

```{r}
linear_reg() %>% 
  set_engine("lm")

```

Try typing `?linear_reg()` in the console to see all available engines and other details about this model type.

Create model specification:

```{r}
lm_mod <- 
  linear_reg() %>% 
  set_engine("lm")
```

Fit model:

```{r}
lm_fit <- 
  lm_mod %>% 
  fit(width ~ initial_volume * food_regime, data = urchins)
lm_fit
```

Present model results in a tidyverse friendly way with `tidy()` from `broom` package.

```{r}
tidy(lm_fit)
```

## [Use a model to predict](https://www.tidymodels.org/start/models/#predict-model)

New example data to predict:

```{r}
new_points <- expand.grid(initial_volume = 20, 
                          food_regime = c("Initial", "Low", "High"))
new_points
```

Generate the mean body width values: 

```{r}
mean_pred <- predict(lm_fit, new_data = new_points)
mean_pred
```

Get confidence intervals and plot:

```{r}
conf_int_pred <- predict(lm_fit, 
                         new_data = new_points, 
                         type = "conf_int")
conf_int_pred

# Now combine: 
plot_data <- 
  new_points %>% 
  bind_cols(mean_pred) %>% 
  bind_cols(conf_int_pred)

# and plot:
ggplot(plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, 
                    ymax = .pred_upper),
                width = .2) + 
  labs(y = "urchin size")
```

## [Model with a different engine](https://www.tidymodels.org/start/models/#new-engine)

Switch to Bayesian approach by simply changing your engine to **stan**: 

```{r}
# set the prior distribution
prior_dist <- rstanarm::student_t(df = 1)

set.seed(123)

# make the parsnip model
bayes_mod <-   
  linear_reg() %>% 
  set_engine("stan", 
             prior_intercept = prior_dist, 
             prior = prior_dist) 

# train the model
bayes_fit <- 
  bayes_mod %>% 
  fit(width ~ initial_volume * food_regime, data = urchins)

print(bayes_fit, digits = 5)
```

To update the parameter table, the `tidy()` method is once again used: 

```{r}
tidy(bayes_fit, intervals = TRUE)
```

Get your predictions without changing the syntax you used earlier:

```{r}
bayes_plot_data <- 
  new_points %>% 
  bind_cols(predict(bayes_fit, new_data = new_points)) %>% 
  bind_cols(predict(bayes_fit, new_data = new_points, type = "conf_int"))

ggplot(bayes_plot_data, aes(x = food_regime)) + 
  geom_point(aes(y = .pred)) + 
  geom_errorbar(aes(ymin = .pred_lower, ymax = .pred_upper), width = .2) + 
  labs(y = "urchin size") + 
  ggtitle("Bayesian model with t(1) prior distribution")
```


Think about how we are using the pipe (`%>%`):

+ Use the pipe to pass around the _data_ in the **tidyverse**   
+ Use the pipe to pass around the _model object_ with **tidymodels**
