---
title: "cloudstart"
output: github_document
---

## R Markdown documents for RStudio Cloud ☁ to accompany tidymodels articles

This repository contains R Markdown documents to accompany the [Get Started](https://www.tidymodels.org/start/) articles on https://www.tidymodels.org/. 

The R Markdown documents in this repo are here to help you walk through these articles by running identical code chunks. Take advantage of the RStudio IDE and use "Run All Chunks Above" or "Run Current Chunk" buttons to easily execute them. 

## Get Started

Two options to get started:

+ You can start by clicking on the R Markdown file of your choice in the Files pane.

+ For a more user friendly setup, press the play button at the top-right of the code chunk provided for each article: 

1. [Build a model](https://www.tidymodels.org/start/models/):

```{r}
viewer <- getOption("viewer")
viewer("https://www.tidymodels.org/start/models/", 
       height = "maximize")
rstudioapi::navigateToFile("01_build_a_model.Rmd")
```

2. [Preprocess your data with recipes](https://www.tidymodels.org/start/recipes):

```{r}
viewer <- getOption("viewer")
viewer("https://www.tidymodels.org/start/recipes/", 
       height = "maximize")
rstudioapi::navigateToFile("02_preprocess_with_recipes.Rmd")
```

3. [Evaluate your model with resampling](https://www.tidymodels.org/start/resampling)  

```{r}
viewer <- getOption("viewer")
viewer("https://www.tidymodels.org/start/resampling/", 
       height = "maximize")
rstudioapi::navigateToFile("03_evaluate_with_resampling.Rmd")
```

4. [Tune model parameters](https://www.tidymodels.org/start/tuning/)  

```{r}
viewer <- getOption("viewer")
viewer("https://www.tidymodels.org/start/tuning/", 
       height = "maximize")
rstudioapi::navigateToFile("04_tune_model_parameters.Rmd")
```

5. [A predictive modeling case study](https://www.tidymodels.org/start/case-study/)

```{r}
viewer <- getOption("viewer")
viewer("https://www.tidymodels.org/start/case-study/", 
       height = "maximize")
rstudioapi::navigateToFile("05_case_study.Rmd")
```


If you ever get lost, you can use the link provided with each section header (for example `## [Introduction](https://www.tidymodels.org/start/models/#intro)`) to visit the accompanying section of the online article.

Feel free to change the code, and explore the tidymodels functions and objects. Happy learning! 🤓